Data Cleaning

### 2025-12-04 Data Cleaning.Rmd

@@

```{r}
library(dplyr)
library(tidyr)

@@

library(readr)
write_csv(merged_clean2, "alkalinity_hardness_cleaned.csv")

@@

library(lubridate)
library(ggplot2)
library(zoo)
library(purrr)

@@

```{r}
## Export CSV of cleaned raw water data

write_csv(merged_daily_data, )
```

>>>>

```{r}
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(ggplot2)
library(zoo)
library(purrr)

@@

@@

write_csv(merged_clean2, "alkalinity_hardness_cleaned.csv")

@@

@@

```{r}
## Export CSV of cleaned raw water data
# run once, only after processing
write_csv(merged_daily_data, "RWAR_temp_pH_cond_turb_cleaned.csv")

## Coagulant and Polymer Doses

doses <- read.csv("CoagulantDoses_2017-20.csv")
doses <- doses |> 
  mutate(Timestamp = mdy(doses$Date)) |>
  select(-c("Date"))

ggplot(doses) +
  geom_point(aes(x=Timestamp, y=Ferric.chloride..mg.L), color="red") +
  geom_point(aes(x=Timestamp, y=Cationic.Polimer..mg.L), color="blue") +
  xlab("Date")+
  ylab("Dosing (mg/L)")+
  ggtitle("Doses over Time (FeCl red, Cationic Polymer blue)")

RWAR_dosing_cleaned <- merge(doses, merged_daily_data, by="Timestamp")

```{r}
## Export to CSV
# only execute after run processing steps above
write_csv(RWAR_dosing_cleaned, "RWAR_dosing_cleaned.csv")

```{r}
water_lm <- lm(Ferric.chloride..mg.L ~ ., 
               data=RWAR_dosing_cleaned |> select(-c(Timestamp, Cationic.Polimer..mg.L)))
summary(water_lm)

```{r}
data <- read.csv("RWAR_dosing_cleaned.csv")
data |>
  mutate(Temp_lag1 = lag(mean_Temp, 1), 
         Temp_lag2 = lag(mean_Temp, 2),
         pH_lag1 = lag(mean_pH, 1),
         pH_lag2 = lag(mean_pH, 2),
         Conductivity_lag1 = lag(conductivity, 1),
         Conductivity_lag2 = lag(conductivity, 2))
